Prestige bias in cultural evolutionary dynamics

### 2023-10-29 ms/ms.tex
continued revision with SE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Adaptive success-bias weight}

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal success-bias weight, $\alpha^*$, which minimizes the expected squared error between the estimated and the ideal trait values,
\begin{equation}
\alpha^* = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose role-model $j$.

We find that when copiers {choose} their success-bias weight, it decreases with the number of copiers that have already chosen a role-model~(\Cref{fig:influence_advantage}).
Moreover, their estimation error is much lower compared to a constant success-bias weight, which gives roughly the same high estimation error to all copiers (compare \Cref{fig:influence_advantage}B and C): in this example, the adaptive weight estimation error converges to $0.046$, whereas a constant weight gives values $>0.74$.

\begin{figure}[h]
    \includegraphics[width=\linewidth]{choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{A}) and estimation error (\textbf{B}) decrease during the role-model choosing process, demonstrating that influence becomes more favored as more copiers have made their choice.
However, when $\alpha$ is homogeneous (\textbf{C}), the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

@@

We developed a cultural-evolution model with prestige bias that includes both indirect success and influence biases, where the latter is a bias towards role-models with many copiers {and hence is the same as conformity bias}.
We model these biases using a stochastic role-model choice process: each copier, in {turn}, randomly chooses a role-model, and this choice is affected both by the estimated success of each potential role-model and the number of copiers that already chose each role-model (eq.~\ref{eq:prestige}).

@@

{Prestige bias also occurs} in more modern domains such as western medicine.
Norredam and Album~\citep{medical_prestige} examined literature from 1950 to 2005 on the effects of prestige on medicinal practices. They found that active, specialized, biomedical, and high-technological types of medicine on organs in the upper part of the bodies of young and middle-aged people were accorded high levels of prestige, whereas medicine and practices that were not of these types had low levels of prestige.
For example, they found that surgery counts as the most prestigious specialty, while psychiatry is the {least} prestigious.
In addition, doctors tend to rank practices that require more time to master as more prestigious, while other procedures that are considered easier to master are less prestigious. 
This means that there may be very important practices that are neglected due to prestige bias.
They concluded that such differences in prestige may {affect} for actual priority setting in healthcare systems.

@@

\bibitem{medical_prestige}
Norredam, Marie, and Dag Album. "Prestige and its significance for medical specialties and diseases." Scandinavian journal of public health 35.6 (2007): 655-661.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Optimal	 success-bias weight}

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal success-bias weight, $\alpha^*$, which minimizes the expected squared error between the chosen trait value and the optimal trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^* = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose role-model~$j$.
% TODO statement about fitness

We find that when copiers choose their success-bias weight according to \cref{eq:optimal_alpha}, it decreases with the number of copiers that have already chosen a role-model~(\Cref{fig:influence_advantage}).
Moreover, their estimation error is much lower compared to a constant success-bias weight, which gives roughly the same high estimation error to all copiers (compare \Cref{fig:influence_advantage}B and C): in this example, the optimal success-bias weight gives an estimation error (difference between the chosen and optimal trait) converges to $0.046$, whereas a constant success-bias weight gives values $>0.74$.

\begin{figure}[h]
    \includegraphics[width=\linewidth]{choose_bias.pdf}
  \caption{
  \textbf{Advantage of an optimal success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{A}) and estimation error (\textbf{B}) decrease during the role-model choosing process (within a single generation), demonstrating that influence becomes more favored by copiers as more copiers have made their choice.
However, when $\alpha$ is homogeneous (\textbf{C}), the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the optimal $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

@@

We developed a cultural-evolution model with prestige bias that includes both indirect success and influence biases, where the latter is a bias towards role-models with many copiers.
We model these biases using a stochastic role-model choice process: each copier, in {turn}, randomly chooses a role-model, and this choice is affected both by the estimated success of each potential role-model and the number of copiers that already chose each role-model (eq.~\ref{eq:prestige}).

@@

@@
